Viscous Transport in Eroding Porous Media

### 2019-09-18 docs/transport/erosion_transport.tex
edits up to the numerical results

The erosion rate loses differentiability if the shear stress changes
sign, and this leads to corners developing on $\gamma$ and numerical
instabilities.  Therefore, we modify the erosion rate, $V_\nn$ in
equation~\eqref{eqn:erosionModel}, with
\begin{align}
  \Vn = \CE \, \abs{\tau} + \epsilon \langle\abs{\tau}\rangle \left(
    \frac{L}{2\pi} \kappa - 1 \right),
\end{align}
where $\epsilon \ll 1,$ $\langle \cdot \rangle$ is the spatial average,
$L$ is the length of $\gamma$, and $\kappa$ is the curvature of
$\gamma$.  This new term penalizations regions of high curvature, and
the normalization term guarantees that the regularization only changes
the body's shape, but not its total length.  Finally, to increase the
overall stability of the method, a narrow Gaussian filter is applied to
the erosion rate at each time step.

@@

Once~\eqref{eqn:BIE} is solved for the density function $\eeta$, the
Stokeslets $\llambda_\ell$, and the rotlets $\xi_\ell$, the deformation
tensor, pressure, and vorticity of the double-layer potential can all be
computed analytically.  For $\xx \in \Omega$, the deformation tensor,
pressure, and vorticity of the double-layer
potential~\eqref{eqn:velocityDLP} are
\begin{align}
  \label{eqn:deformationDLP} 
  \ssigma(\xx) &= \int_{\bd\Omega} \mathbf{D}_{\ssigma}(\xx,\yy) 
      \eeta(\yy) ds_{\yy} 
  = \frac{1}{2\pi}\int_{\bd\Omega} \left(
    2\frac{\rr \cdot \nn}{\rho^2} \frac{\rr \cdot \eeta}{\rho^2} \II + 
    \frac{\rr \cdot \eeta}{\rho^4} (\nn \otimes \rr + \rr \otimes \nn) 
    \right. \nonumber \\
    & \qquad\qquad\qquad\qquad\qquad\qquad \left.
    +\frac{\rr \cdot \nn}{\rho^4} (\eeta \otimes \rr + \rr \otimes \eeta) - 
    8\frac{(\rr \cdot \nn)(\rr \cdot \eeta)}{\rho^6}(\rr \otimes \rr)
    \right) \, ds_\yy, \\
%
  p(\xx) &= \int_{\bd\Omega} D_{p}(\xx,\yy) 
      \eeta(\yy) ds_{\yy} 
  = -\frac{1}{\pi}\int_{\bd\Omega} \frac{1}{\rho^2}
    \left(I - 2 \frac{\rr \otimes \rr}{\rho^2}\right) \nn \cdot
      \eeta(\yy)\,ds_\yy, \\
%
  \label{eqn:vorticityDLP} 
  \omega(\xx) &= \int_{\bd\Omega} D_{\omega}(\xx,\yy) 
      \eeta(\yy) ds_{\yy} 
  = -\frac{1}{\pi} \int_{\bd\Omega}
    \frac{(\rr \cdot \nn^\perp) + (\rr \cdot \nn)}{\rho^4} 
      (\rr \cdot \eeta) \, ds_\yy,
\end{align}
respectively. We require the deformation tensor of the double-layer potential at
$\xx \in \bd \Omega$.  Since the deformation tensor of the double-layer
potential is discontinuous across $\bd\Omega$, the jump term 
\begin{align}
  \frac{1}{2} \left(\pderiv{\eeta}{\ss} \cdot \ss \right) \left[
    \begin{array}{cc}
      s_x^2 - s_y^2 & 2s_x s_y \\ 2s_x s_y & s_y^2 - s_x^2
    \end{array}
  \right],
  \label{eqn:deformationJump}
\end{align}
must be added to $\ssigma(\xx)$ for $\xx \in \bd\Omega$.  Finally, the
Stokeslets and rotlets do not have any jumps on $\gamma$, so the
deformation tensor, pressure, and vorticity due to the Stokeslets and
rotlets are straightforward to compute~\cite{poz1992} \todo[inline]{BQ:
Make sure this is the right reference}.  Having computed the deformation
tensor on $\gamma$, the shear stress is computed as defined in
equation~\eqref{eqn:shearStress}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Cauchy Integral Representation of the Double-Layer
Potential}
\label{sec:DLPcomplex}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
In equations~\eqref{eqn:velocityDLP} and~\eqref{eqn:deformationDLP},
$\uu$ and its deformation tensor are written as a contour integral in
$\RR^2$.  However, so that we can apply a particular quadrature method,
we write these layer potentials as Cauchy integrals in
$\CC$ and their derivatives.  The first step is to write the Laplace
double-layer potential as the complex integral
\begin{align}
  \DD[\eta](\xx) = \frac{1}{2\pi} \int_{\bd\Omega} 
    \frac{\rr \cdot \nn}{\rho^2}\, ds_\yy = \Re (v(x)) \quad 
    \text{where} \quad v(x) = \frac{1}{2\pi i} \int_{\bd\Omega}
    \frac{\eta(y)}{x - y} \, dy,
  \label{eqn:laplaceComplex}
\end{align}
where $\xx,\yy \in \RR^2$ are identified with the points $x,y \in \CC$.
Equation~\eqref{eqn:laplaceComplex} is converted to a Cauchy
integral by first finding the boundary data of $v$.  Assuming that
$\Omega$ is a simply-connected interior domain, the boundary data
of $v$ satisfies the Sokhotski-Plemelj jump relation
\begin{align}
  \label{eqn:SPrelation}
  v(x) = - \frac{1}{2} \eta(x) + \frac{1}{2\pi i} \int_{\bd\Omega}
    \frac{\eta(y)}{x-y}\, dy, \quad x \in \bd\Omega.
\end{align}
For exterior domains, the jump term changes from $-1/2$ to $1/2$, and
for multiply-connected domains, such as a porous media, $\bd\Omega$ is
simply decomposed into its different connected components and the
appropriate Cauchy integral is evaluated.  Since $v$ is a holomorphic
function and we have its boundary data, by the Cauchy integral theorem
we have
\begin{align}
  \label{eqn:cauchy}
  v(x) = \frac{1}{2\pi i}\int_{\bd\Omega} 
    \frac{v(y)}{y-x} \,dy, \quad
  v'(x) = \frac{1}{2\pi i} \int_{\bd\Omega}
    \frac{v(y)}{(y-x)^2} \, dy, \quad
  v''(x) = \frac{1}{\pi i} \int_{\bd\Omega}
    \frac{v(y)}{(y-x)^3} \, dy,
\end{align}
for $x \in \Omega$.  Since $v(x)$ depends on the density function $\eta
\in \CC$, we use the notation $v[\eta](x)$ to denote the holomorphic
function defined in equation~\eqref{eqn:laplaceComplex}, and its first
two derivatives are written as $v'[\eta](x)$ and $v''[\eta](x)$.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Anomalous Dispersion}
\label{sec:dispersion}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The dynamics of fluid in a porous media is often characterized in terms
of anomalous dispersion~\cite{kla-rad-sok2008, den-cor-sch-ber2004}.
The anomalous dispersion rate depends on the porosity and
permeability~\cite{koc-bra1988}, but it is also affected by the
distribution and shape of the grains. We calculate the anomalous
dispersion rates by analyzing the streamlines governed by
equation~\eqref{eqn:tracers} in eroded geometries. Given a set of $N_p$
trajectories, we define $\lambda_j(t)$ to be the arclength of the
trajectory
\begin{align}
  \lambda_j(t) = \int_{0}^t \|\ss'_j(\tau)\|\, d\tau, 
    \quad j=1,\ldots,N_p,
\end{align}
and the first and second ensemble moments
\begin{align}
  \label{eqn:moments}
  \langle \lambda \rangle (t) = 
    \frac{1}{N_p} \sum_{j=1}^{N_p} \lambda_j(t), \quad 
    \sigma_\lambda^2(t) = \frac{1}{N_p} 
    \left[\lambda_j(t) - \langle \lambda \rangle(t) \right]^2.
\end{align}
Then,
particle dispersion is characterized as $\sigma_\lambda$.  At early
times, the particles have not explored much of the geometry, and we
expect a ballistic motion $\sigma_\lambda \sim t$.  However, as the
particles pass the grains, their trajectories are altered, and we expect
that $\sigma_\lambda \sim t^\alpha$, with $\alpha \in (0.5,1)$,
indicating that the flow is super-dispersive.

To establish an anomalous dispersion rate, the trajectories must pass
several grains so that an asymptotic regime is reached. The geometries
that we consider are too short to observe the asymptotic dispersion, so
we use a reinsertion method to form longer trajectories. Similar to the
work of others~\cite{dea-qua-bir-jua2018, puy-gou-den2019}, once a
particle reaches the end of the porous region, it is reinserted at the
start of the porous region.  To minimize errors caused by the
reinsertion, the particle is initialized at the discretization point
that has the closest velocity to the particles velocity at the outlet.
After a single trajectory is formed, it has undergone a collection of
reinsertions.  The fore, as a post-processing step, the trajectory is made
continuous by attaching the tail of each segment to the origin of the
next segment.

We consider trajectories in an open channel so that they can be compared
to trajectories in eroded geometries with large permeability.  The
trajectory originating at $(s_1(0),s_2(0))$ is
\begin{align}
  s_1(t) = (1-s_2(0)^2)t + s_1(0), \quad
  s_2(t) = s_2(0).
\end{align}
Averaging over $s_2(0) \in [-1,1]$, the first and second ensemble
moments are
\begin{align}
  \langle \lambda \rangle (t) = \frac{2}{3}t, \quad 
    \sigma_\lambda^2(t) = \frac{4}{45}t^2,
\end{align}
so the particle motion is ballistic since $\sigma_\lambda \sim t$.  In
Section~\ref{sec:results}, we compare these moments in an open channel
with the trajectories in an eroded geometry.

@@

The choice for the trajectory with length $\lambda(y)$ is not universal,
and Duda et al.~\cite{dud-koz-mat2011} summarize several options. One
choice, which we adopt, considers trajectories that originate at $S$ and
terminate at a parallel cross-section.  We use the cross-sections $x=-1$
and $x=1$.  This choice of averaging allows us to also compute the
tortuosity in terms of an area integral.  Under the assumptions that the
flow is incompressible and not re-entrant, meaning that all streamlines
connect the two cross-sections, the tortuosity in
equation~\eqref{eqn:tortuosity1} is equivalent to~\cite{dud-koz-mat2011}
\begin{align}
  T = \frac{\displaystyle\int_\Omega \|\uu(\yy)\|\, d\yy}
           {\displaystyle\int_\Omega u_x(\yy)\, d\yy},
  \label{eqn:tortuosity2}
\end{align}
where $\Omega$ is the fluid region between the inlet and outlet
cross-sections.  Recirculation zones are possible in viscous
fluids~\cite{hig1985}, but they are very small in the examples we
consider and have a negligible effect on the tortuosity.  Since
equation~\eqref{eqn:tortuosity2} does not require the additional work of
computing particle trajectories at every time step, we use this
definition for the majority of the tortuosity calculations.  However, we
do compare the two formulas for the tortuosity at several time steps in
Section~\ref{sec:results}.

There have been efforts to relate the tortuosity to the porosity.  For
example, Matyka et al.~\cite{mat-kha-koz2008} propose the models
\begin{subequations}
  \label{eqn:tortuosityModels}
  \begin{align}
    \widehat{T}(\phi) &= \phi^{-p}, \\
    \widehat{T}(\phi) &= 1-p \log \phi, \\
    \widehat{T}(\phi) &= 1+p (1-\phi), \\
    \widehat{T}(\phi) &= (1+p (1-\phi))^2, 
  \end{align}
\end{subequations}
where $p>0$ is a fitting parameter.  In Section~\ref{sec:results}, we
compare these four models with the tortuosity of eroding geometries.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Pore Throat Size}
\label{sec:throats}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{itemize}
  \item Pore throat sizes affect many things such as first passage time,
  mixing, chemical reactions, heat conduction, biological activities
\end{itemize}

\begin{figure}[H]
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[height =0.8\linewidth]{./figs/triangulation_100b100}
\caption{}
\end{subfigure}
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[height = 0.8\linewidth]{./figs/gap_hist100b100}
\caption{}
\end{subfigure}
\caption{\label{fig:Eroding100gap} (a): The Delaunay triangulation of the porous media. 
(b): The distribution of gaps from (a) with the fitting curve of Weibull distribution.}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Numerical Methods}
\label{sec:method}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
In line with our previous work~\cite{qua-moo2018}, we use two meshes to
simulate erosion. The integral equation is solved by discretizing the
boundary of the geometry at a set of points distributed equally in
arclength, and then applying a collocation method
(Section~\ref{sec:spatialDiscretization}).  The quadrature rule is
either the trapezoid rule or the Barycentric quadrature rule
(Section~\ref{sec:bary}).  The selection algorithm that determines which
quadrature method is applied is described in Section~\ref{sec:fmm}.
Once the shear stress is computed, the bodies are eroded a single time
step by using a {\thL} discretization~\cite{hou-low-she1994}.  By using
a {\thL} formulation, the stiff curvature term is linear and the
discretization points remain equispaced in arclength.  Finally, time
stepping methods for the eroding bodies and passive particles are
described in Section~\ref{sec:time}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Spatial Discretization}
\label{sec:spatialDiscretization}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
By using an integral equation formulation, we only have to discretize
the one-dimensional boundary of the domain.  Therefore, we easily
resolve complex geometries, and differentiation is performed with
spectral accuracy in Fourier space.  We discretize each eroding body
$\gamma_k$ with $N_\iin$ points and discretize the outer wall $\Gamma$
with $N_\out$ points.  The $j^{th}$ discretization point on $\Gamma$ and
$\gamma_k$ are denoted by $\yy_j^0$ and $\yy^k_j$, respectively.  The
discretization points are initially distributed evenly in arclength, and
this equispacing is maintained for the entire simulation by using the
{\thL} formulation.  In addition, we apply mild smoothing
strategies~\cite{qua-moo2018} to avoid the corners that erosion
inevitably develop.

>>>>

@@

Once~\eqref{eqn:BIE} is solved for the density function $\eeta$, the
Stokeslets $\llambda_\ell$, and the rotlets $\xi_\ell$, the deformation
tensor, pressure, and vorticity of the double-layer potential can all be
computed analytically.  For $\xx \in \Omega$, the deformation tensor,
pressure, and vorticity of the double-layer
potential~\eqref{eqn:velocityDLP} are
\begin{align}
  \label{eqn:deformationDLP} 
  \ssigma(\xx) &= \int_{\bd\Omega} \mathbf{D}_{\ssigma}(\xx,\yy) 
      \eeta(\yy) ds_{\yy} 
  = \frac{1}{2\pi}\int_{\bd\Omega} \left(
    2\frac{\rr \cdot \nn}{\rho^2} \frac{\rr \cdot \eeta}{\rho^2} \II + 
    \frac{\rr \cdot \eeta}{\rho^4} (\nn \otimes \rr + \rr \otimes \nn) 
    \right. \nonumber \\
    & \qquad\qquad\qquad\qquad\qquad\qquad \left.
    +\frac{\rr \cdot \nn}{\rho^4} (\eeta \otimes \rr + \rr \otimes \eeta) - 
    8\frac{(\rr \cdot \nn)(\rr \cdot \eeta)}{\rho^6}(\rr \otimes \rr)
    \right) \, ds_\yy, \\
%
  p(\xx) &= \int_{\bd\Omega} D_{p}(\xx,\yy) 
      \eeta(\yy) ds_{\yy} 
  = -\frac{1}{\pi}\int_{\bd\Omega} \frac{1}{\rho^2}
    \left(I - 2 \frac{\rr \otimes \rr}{\rho^2}\right) \nn \cdot
      \eeta(\yy)\,ds_\yy, \\
%
  \label{eqn:vorticityDLP} 
  \omega(\xx) &= \int_{\bd\Omega} D_{\omega}(\xx,\yy) 
      \eeta(\yy) ds_{\yy} 
  = -\frac{1}{\pi} \int_{\bd\Omega}
    \frac{(\rr \cdot \nn^\perp) + (\rr \cdot \nn)}{\rho^4} 
      (\rr \cdot \eeta) \, ds_\yy,
\end{align}
respectively. We require the deformation tensor of the double-layer potential at
$\xx \in \bd \Omega$.  Since the deformation tensor of the double-layer
potential is discontinuous across $\bd\Omega$, the jump term 
\begin{align}
  \frac{1}{2} \left(\pderiv{\eeta}{\ss} \cdot \ss \right) \left[
    \begin{array}{cc}
      s_x^2 - s_y^2 & 2s_x s_y \\ 2s_x s_y & s_y^2 - s_x^2
    \end{array}
  \right],
  \label{eqn:deformationJump}
\end{align}
must be added to $\ssigma(\xx)$ for $\xx \in \bd\Omega$.  Finally, the
Stokeslets and rotlets do not have any jumps on $\gamma$, so the
deformation tensor, pressure, and vorticity due to the Stokeslets and
rotlets are straightforward to compute~\cite{poz1992}.  Having computed
the deformation tensor on $\gamma$, the shear stress is computed as
defined in equation~\eqref{eqn:shearStress}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Cauchy Integral Representation of the Double-Layer
Potential}
\label{sec:DLPcomplex}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
In equations~\eqref{eqn:velocityDLP} and~\eqref{eqn:deformationDLP},
$\uu$ and its deformation tensor are written as a contour integral in
$\RR^2$.  However, so that we can apply a particular quadrature method,
we write these layer potentials as Cauchy integrals in
$\CC$ and their derivatives.  The first step is to write the Laplace
double-layer potential as the complex integral
\begin{align}
  \DD[\eeta](\xx) = \frac{1}{2\pi} \int_{\bd\Omega} 
    \frac{\rr \cdot \nn}{\rho^2}\eeta(\yy)\, ds_\yy = \Re (v(x)) \quad 
    \text{where} \quad v(x) = \frac{1}{2\pi i} \int_{\bd\Omega}
    \frac{\eta(y)}{x - y} \, dy,
  \label{eqn:laplaceComplex}
\end{align}
where $\xx,\yy \in \RR^2$ are identified with the points $x,y \in \CC$.
Equation~\eqref{eqn:laplaceComplex} is converted to a Cauchy
integral by first finding the boundary data of $v$.  Assuming that
$\Omega$ is a simply-connected interior domain, the boundary data
of $v$ satisfies the Sokhotski-Plemelj jump relation
\begin{align}
  \label{eqn:SPrelation}
  v(x) = - \frac{1}{2} \eta(x) + \frac{1}{2\pi i} \int_{\bd\Omega}
    \frac{\eta(y)}{x-y}\, dy, \quad x \in \bd\Omega.
\end{align}
For exterior domains, the jump term changes from $-1/2$ to $1/2$, and
for multiply-connected domains, such as a porous media, $\bd\Omega$ is
simply decomposed into its different connected components and the
appropriate Cauchy integral is evaluated.  Since $v$ is a holomorphic
function and we have its boundary data, by the Cauchy integral theorem
we have
\begin{align}
  \label{eqn:cauchy}
  v(x) = \frac{1}{2\pi i}\int_{\bd\Omega} 
    \frac{v(y)}{y-x} \,dy, \quad
  v'(x) = \frac{1}{2\pi i} \int_{\bd\Omega}
    \frac{v(y)}{(y-x)^2} \, dy, \quad
  v''(x) = \frac{1}{\pi i} \int_{\bd\Omega}
    \frac{v(y)}{(y-x)^3} \, dy,
\end{align}
for $x \in \Omega$.  Since $v(x)$ depends on the density function $\eta
\in \CC$, we use the notation $v[\eta](x)$ to denote the holomorphic
function defined in equation~\eqref{eqn:laplaceComplex}, and its first
two derivatives are written as $v'[\eta](x)$ and $v''[\eta](x)$.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Anomalous Dispersion}
\label{sec:dispersion}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The dynamics of fluid in a porous media is often characterized in terms
of anomalous dispersion~\cite{kla-rad-sok2008, den-cor-sch-ber2004}.
The anomalous dispersion rate depends on the porosity and
permeability~\cite{koc-bra1988}, but it is also affected by the
distribution and shape of the grains. We calculate the anomalous
dispersion rates by analyzing the streamlines governed by
equation~\eqref{eqn:tracers} in eroded geometries. Given a set of $N_p$
trajectories, we define $\lambda_j(t)$ to be the arclength of the
trajectory
\begin{align}
  \lambda_j(t) = \int_{0}^t \|\ss'_j(\tau)\|\, d\tau, 
    \quad j=1,\ldots,N_p,
\end{align}
and the first and second ensemble moments
\begin{align}
  \label{eqn:moments}
  \langle \lambda \rangle (t) = 
    \frac{1}{N_p} \sum_{j=1}^{N_p} \lambda_j(t), \quad 
    \sigma_\lambda^2(t) = \frac{1}{N_p} 
    \left[\lambda_j(t) - \langle \lambda \rangle(t) \right]^2.
\end{align}
Particle dispersion is characterized as $\sigma_\lambda$.  At early
times, the particles have not explored much of the geometry, and we
expect a ballistic motion $\sigma_\lambda \sim t$.  However, as the
particles pass the grains, their trajectories are altered, and we expect
that $\sigma_\lambda \sim t^\alpha$, with $\alpha \in (0.5,1)$,
indicating that the flow is super-dispersive.

To establish an anomalous dispersion rate, the trajectories must pass
several grains so that an asymptotic regime is reached. The geometries
that we consider are too short to observe the asymptotic dispersion, so
we use a reinsertion method to form longer trajectories. Similar to the
work of others~\cite{dea-qua-bir-jua2018, puy-gou-den2019}, once a
particle reaches the end of the porous region, it is reinserted at the
start of the porous region.  To minimize errors caused by reinsertion,
the particle is initialized at the discretization point that has the
closest velocity to the particle's velocity at the outlet.  After a
single trajectory is formed, it has undergone a collection of
reinsertions.  Therefore, as a post-processing step, the trajectory is
made continuous by attaching the tail of each segment to the origin of
the next segment.

We consider trajectories in an open channel so that they can be compared
to trajectories in eroded geometries with large permeability.  The
trajectory originating at $(s_1(0),s_2(0))$ is
\begin{align}
  s_1(t) = (1-s_2(0)^2)t + s_1(0), \quad
  s_2(t) = s_2(0).
\end{align}
Averaging over $s_2(0) \in [-1,1]$, the first and second ensemble
moments are
\begin{align}
  \langle \lambda \rangle (t) = \frac{2}{3}t, \quad 
    \sigma_\lambda^2(t) = \frac{4}{45}t^2.
\end{align}
In Section~\ref{sec:results}, we compare these moments in an open
channel with the trajectories in an eroded geometry.

@@

The definition for the tortuosity depends on how the streamlines are
defined, and Duda et al.~\cite{dud-koz-mat2011} summarize several
options. One choice, that we adopt, considers streamlines that originate
at $S$ and terminate at a parallel cross-section.  We use the
cross-sections $x=-1$ and $x=1$.  This choice of averaging allows us to
also compute the tortuosity in terms of an area integral.  Under the
assumptions that the flow is incompressible and not re-entrant, meaning
that all streamlines connect the two cross-sections, the tortuosity in
equation~\eqref{eqn:tortuosity1} is equivalent to~\cite{dud-koz-mat2011}
\begin{align}
  T = \frac{\displaystyle\int_\Omega \|\uu(\yy)\|\, d\yy}
           {\displaystyle\int_\Omega u_x(\yy)\, d\yy},
  \label{eqn:tortuosity2}
\end{align}
where $\Omega$ is the fluid region between the inlet and outlet
cross-sections.  Recirculation zones are possible in viscous
fluids~\cite{hig1985}, but they are very small in the examples we
consider and have a negligible effect on the tortuosity.  Since
equation~\eqref{eqn:tortuosity2} does not require the additional work of
computing particle trajectories at every time step, we use this
definition for the majority of the tortuosity calculations.  However, we
do compare the two definitions for the tortuosity at several time steps
in Section~\ref{sec:results}.

There have been efforts to relate the tortuosity to the porosity.  For
example, Matyka et al.~\cite{mat-kha-koz2008} propose the models
\begin{subequations}
  \label{eqn:tortuosityModels}
  \begin{align}
    \widehat{T}(\phi) &= \phi^{-p}, \\
    \widehat{T}(\phi) &= 1-p \log \phi, \\
    \widehat{T}(\phi) &= 1+p (1-\phi), \\
    \widehat{T}(\phi) &= (1+p (1-\phi))^2, 
  \end{align}
\end{subequations}
where $\phi$ is the porosity and $p>0$ is a fitting parameter.  In
Section~\ref{sec:results}, we compare these four models with the
tortuosity of eroding geometries.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Pore Throat Size}
\label{sec:throats}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
While transport in a porous media depends on the porosity, it also
depends on the placement of the grains.  In particular, the grain
placement affects velocity scales~\cite{ali-par-wei-bre2017},
correlation structures~\cite{leb-ded-dav-bou2007}, contaminant
transport~\cite{knu-car2005},
channelization~\cite{sie-ili-pri-riv-gua2019}, and more. In addition,
pore network models~\cite{bry-kin-mel1993, bry-mel-cad1993, bij-blu2006}
rely on descriptions of the pore throat sizes. To characterize the grain
placement, we compute distributions of gap sizes between neighboring
grains.  To define neighboring grains, we form a Delaunay triangulation
of nodes placed at the center of each eroding grain and at a collection
of points around the boundary of the porous region.  Then, we say that
two grains are neighbors if their centers share an edge of the
triangulation~\cite{dea-qua-bir-jua2018}.  Finally, we compute the
minimum distance between all pairs of neighboring grains.  In
Figure~\ref{fig:Eroding100gap}(a), we plot a histogram of the pore size
of 100 eroded bodies.  The black curve is an un-normalized Weibull
distribution with the same mean and variance as the measured data.  In
Figure~\ref{fig:Eroding100gap}(b), we plot lines connected the two
closest points of all neighboring gaps. We do not consider pores between
eroding bodies and the solid and the artificial points placed around the
porous region, so some of the grains near the porous region boundary
only have two neighbors.

\begin{figure}[H]
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[height = 0.8\linewidth]{./figs/gap_hist100b100}
\caption{}
\end{subfigure}
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[height =0.8\linewidth]{./figs/triangulation_100b100}
\caption{}
\end{subfigure}
\caption{\label{fig:Eroding100gap} (a): The distribution of the pore
sizes of an eroded geometry. The pores are defined as the space between
neighboring grains.  The black curve is the Weibull distribution with
the same first two moments as the data. (b): Two grains are neighbors if
they share a edge in a Delaunay triangulation of the centers of the
grains.}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Numerical Methods}
\label{sec:method}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
In line with our previous work~\cite{qua-moo2018}, we use two meshes to
simulate erosion. The integral equation is solved by discretizing the
boundary of the geometry at a set of points distributed equally in
arclength, and then applying a collocation method
(Section~\ref{sec:spatialDiscretization}).  Depending on the proximity
of the target point to the source points, the quadrature rule is either
the trapezoid rule or the Barycentric quadrature rule
(Section~\ref{sec:bary}).  The selection algorithm that determines which
quadrature method is applied is described in Section~\ref{sec:fmm}.
Once the shear stress is computed, the bodies are eroded a single time
step by using a {\thL} discretization~\cite{hou-low-she1994}.  By using
a {\thL} formulation, the stiff curvature term is linear, treated
implicitly, and the discretization points remain equispaced in
arclength.  Finally, time stepping methods for the eroding bodies and
passive particles are described in Section~\ref{sec:time}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Spatial Discretization}
\label{sec:spatialDiscretization}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
By using an integral equation formulation, we only have to discretize
the one-dimensional boundary of the domain.  Therefore, we easily
resolve complex geometries, and differentiation is performed with
spectral accuracy in Fourier space.  We discretize each eroding body
$\gamma_k$ with $N_\iin$ points and discretize the outer wall $\Gamma$
with $N_\out$ points.  The $j^{th}$ discretization point on $\Gamma$ and
$\gamma_k$ are denoted by $\yy_j^0$ and $\yy^k_j$, respectively.  The
discretization points are initially distributed evenly in arclength, and
this equispacing is maintained for the entire simulation by using the
{\thL} formulation.  In addition, we apply mild smoothing
strategies~\cite{qua-moo2018} to avoid the corners that inevitably
develop in erosion.

Given the discretization points of $\bd\Omega$, applying the trapezoid
rule results in the collocation method for~\eqref{eqn:BIE} 
\begin{subequations}
\label{eqn:trapLinearSystem}
  \begin{alignat}{3}
  \UU_\ell &= \sum_{j=1}^{N_\out} 
    w^0_j D(\yy^0_\ell,\yy^0_j) \eeta_j^0 +
  \sum_{k=1}^M \sum_{j=1}^{N_\iin}
    w^k_j D(\yy^0_\ell,\yy^k_j) \eeta^k_j +
  \sum_{j=1}^{N_\out} w^0_j N_0(\yy^0_\ell,\yy^0_j)\eeta_j^0 
    \nonumber \\
  &\quad+\sum_{k=1}^M S[\llambda_k,\cc_k](\yy^0_\ell) + 
  \sum_{k=1}^M R[\xi_k,\cc_k](\yy^0_\ell),
    &&\hspace{-60pt} \ell = 1,\ldots,N_\out, \\
%
  \mathbf{0} &= \sum_{j=1}^{N_\out} 
    w^0_j D(\yy^m_\ell,\yy^0_j) \eeta_j^0 +
  \sum_{k=1}^M \sum_{j=1}^{N_\iin}
    w^k_j D(\yy^m_\ell,\yy^k_j) \eeta^k_j \nonumber \\
  &\quad+\sum_{k=1}^M S[\llambda_k,\cc_k](\yy^m_\ell) + 
  \sum_{k=1}^M R[\xi_k,\cc_k](\yy^m_\ell),
    &&\hspace{-60pt}\ell = 1,\ldots,N_\iin, \: m=1,\ldots,M, \\
%
  \llambda_m &= \frac{1}{2\pi} \sum_{j=1}^{N_\iin} w_j^m \eeta^m_j,
  \quad \xi_m = \frac{1}{2\pi} \sum_{j=1}^{N_\iin} w_j^m
    (\yy^m_j - \cc_m)^\perp \cdot \eeta^m_j,
    &&\hspace{-60pt} m=1,\ldots,M,
\end{alignat}
\end{subequations}
where $w^k_j$ are quadrature weights that depend on $N_\iin$,
$N_\out$, and the lengths of $\gamma^k$, and $D$ is the kernel of the
Stokes double-layer potential defined in
equation~\eqref{eqn:velocityDLP}.  The linear
system~\eqref{eqn:trapLinearSystem} is a well-conditioned second-kind
integral equation that is solved iteratively with GMRES.  If the number
of discretization points is sufficiently large, then the solution
of~\eqref{eqn:trapLinearSystem} is an accurate approximation of the
density function, Stokeslets, and rotlets.   Then, if $\xx \in \Omega$
is sufficiently far from $\bd\Omega$, an approximation of the velocity,
deformation tensor, and vorticity due to the double-layer potential is
found by applying the trapezoid rule to \eqref{eqn:velocityDLP},
\eqref{eqn:deformationDLP}, and~\eqref{eqn:vorticityDLP}, respectively.
In particular,
\begin{align}
  \label{eqn:trap}
  \uu(\xx) &= \sum_{j=1}^{N_\out} w^0_j D(\xx,\yy^0_j) \eeta_j +
  \sum_{k=1}^M \sum_{j=1}^{N_\iin} w^k_j D(\xx,\yy^k_j) \eeta^k_j, \\
%
  \ssigma(\xx) &= \sum_{j=1}^{N_\out} w^0_j \mathbf{D}_\ssigma(\xx,\yy^0_j) \eeta_j +
  \sum_{k=1}^M \sum_{j=1}^{N_\iin} w^k_j \mathbf{D}_\ssigma(\xx,\yy^k_j)
  \eeta^k_j, \\
%
  w(\xx) &= \sum_{j=1}^{N_\out} w^0_j D_\omega(\xx,\yy^0_j) \eeta_j +
  \sum_{k=1}^M \sum_{j=1}^{N_\iin} w^k_j D_\omega(\xx,\yy^k_j)
  \eeta^k_j.
\end{align}
The contributions due to the Stokeslets and rotlets require no
quadrature and are easily included in the velocity, deformation tensor,
and vorticity.  Finally, to compute the shear stress, we apply the
trapezoid rule with Fourier differentiation to approximate the jump
term~\eqref{eqn:deformationJump} of the deformation tensor, and then the
tensor is applied to the normal and tangent vectors as defined in
equation~\eqref{eqn:shearStress}.
